Using traffic analysis to identify anonymous networks

### 2010-04-26 conference-paper/proposal.tex
First draft of conference submitted

Onion routing was first described in a patent by \cite{Michael:2001}. Like the mix network is uses public key cryptography to provide anonymity, it differs in that messages are repeatedly encrypted then passed through several nodes in a network known as onion routers. Each of these routers is capable of removing a single layer of encryption to reveal delivery instructions. This layering prevents individual nodes obtaining the identity of the sender, receiver and the contents of the message.

@@

The Second Generation Onion Router (Tor) uses a form of Onion Routing 
The Tor network, like the Mix-Net proposed by \textcite{Chaum:1981p296} uses public key encryption and cascades of servers to ensure message anonymity \parencite{Dingledine:2004p314}. In Tor these mixes are known as ‘relays’. Relays that allow connections directly from Tor clients are known as ‘bridges’ and relays that deliver messages from the frontier of the Tor network are known as ‘exit nodes’. Figure \ref{tor-network} shows how a client connects to the Tor network.

\begin{figure}[H]
\center \includegraphics{tor-network-diagram}
\caption{The Tor Network}
\label{tor-network}
\end{figure}

@@

\subsubsection{Expectation-Maximisation}

The first use of machine learning to categorise traffic flows appears in \textcite{McGregor:2004p3826}. A detailed analysis of the attributes that can be used for machine learning and an attempt at coarse grained classification using an Expectation-Maximisation (EM) algorithm are demonstrated. The same technique is also demonstrated in \textcite{Soule:2004p3817} using histograms for finer grained classification. The EM algorithm is again used in \textcite{Zander:2005p6212} and \textcite{Erman:2006p3825}, with the latter also comparing this algorithm favourably against a Naïve Bayes classifier.

\subsubsection{Naïve Bayes}

\textcite{Moore:2005p3827} use a supervised Naïve Bayes to classify traffic flows that have previously been sorted into groups by analysing the flow content. This paper focuses on many of the most commonly used Internet protocols while \textcite{Bonfiglio:2007p6453} uses the technique for identifying traffic belonging to the commercial Skype application.

\textcite{Herrmann:2009p1189} use Bayesian networks to fingerprint visited websites accessed through Privacy Enhancing Technologies (PET), including Tor. This technique performed poorly when applied to the Tor network, but it suggests that Tor traffic has particular characteristics that distinguish it from many existing PETs. It makes a particularly useful observation: “The most frequent packet sizes in the Tor traffic dumps are, in descending order, 1500, −52, −638, 52, 638 and 1150 bytes, accounting for 87.6 \% of all Tor packets.”

\subsubsection{Hidden Markov Model}

Hidden Markov Models are first used as a traffic analysis technique in HMM profiles for network traffic classification \parencite{Wright:2004p3860}. The primary identification characteristics for use with this algorithm are packet size and inter-packet arrival times. With refinement this algorithm is used with increasing accuracy in \textcite{Wright:2006p322} and \textcite{Dainotti:2008p1435}. HMMs are also used in \textcite{Bernaille:2005p6205} to discover distinguishing characteristics of traffic flows, rather than specifically as a classifier.

\subsubsection{Clustering}

@@

\subsubsection{Other Techniques}

@@

\subsubsection{Comparing Techniques}

It is difficult to say what machine learning technique is the most effective as no consensus has been reached, the literature covers a wide variety of techniques each with vastly different goals and no two techniques can be directly compared as the data used for analysis has not been disclosed \parencite{Kim:2007p3867}. However some attempt has been made at comparison in \textcite{Williams:2006p3849} which suggests that the C4.5 algorithm has the greatest performance and accuracy when compared to a number of Bayesian algorithms. \textcite{Mohd:2009p6484} compares thirty machine learning algorithms  to find random tree, IBI, IBK and random forest algorithms obtaining the greatest classification accuracy.

An excellent meta study on traffic analysis techniques is presented in \textcite{Nguyen:2008p3837}. This paper compares many of the published papers, the techniques used, criteria analysed and how effectively they meet stated goals.

@@

\begin{figure}
\center \includegraphics[scale=0.7]{network-diagram}
\caption{Network Diagram}
\label{network-diagram}
\end{figure}

>>>>

Onion routing was first described in a patent by \cite{Michael:2001}. Like the mix network it uses public key cryptography to provide anonymity, it differs in that messages are repeatedly encrypted then passed through several nodes in a network known as onion routers. Each of these routers is capable of removing a single layer of encryption to reveal delivery instructions. This layering prevents individual nodes obtaining the identity of the sender, receiver and the contents of the message.

@@

The Second Generation Onion Router (Tor) uses a form of Onion Routing which employs telescoping circuits to provide low latency anonymity \parencite{Dingledine:2004p314}. A connection through the Tor network constitutes a \emph{circuit}, which hops over several Tor nodes known as \emph{relays}. Relays that allow connections directly from Tor clients are known as \emph{bridges} and relays that deliver messages from the frontier of the Tor network are known as \emph{exit nodes}. Figure \ref{tor-network} shows how a client connects to the Tor network.

\begin{figure}[H]
\center \includegraphics[scale=0.8]{tor-network-diagram}
\caption{The Tor Network}
\label{tor-network}
\end{figure}

@@

There are a vast number of traffic analysis papers that utilize machine learning techniques for the purpose of traffic classification, demonstrating a wide variety of algorithms for this purpose. The first use of machine learning to categorise traffic flows appears in \textcite{McGregor:2004p3826}. A detailed analysis of the attributes that can be used for machine learning and an attempt at coarse grained classification using an Expectation-Maximisation (EM) algorithm are demonstrated. The same technique is also employed in \textcite{Soule:2004p3817} using histograms for finer grained classification. The EM algorithm is again used in \textcite{Zander:2005p6212} and \textcite{Erman:2006p3825}, with the latter also comparing this algorithm favourably against a Na\"ive Bayes classifier.

\textcite{Moore:2005p3827} demonstrates the usage of a supervised Na\"ive Bayes algorithm to classify traffic flows. This paper focuses on many of the most commonly used Internet protocols while \textcite{Bonfiglio:2007p6453} uses the technique for identifying traffic belonging to the commercial Skype application.

\textcite{Herrmann:2009p1189} uses Bayesian networks to fingerprint visited websites accessed through Privacy Enhancing Technologies (PET), including Tor. This technique performed poorly when applied to the Tor network, but it suggests that Tor traffic has particular characteristics that distinguish it from many existing PETs. It makes a particularly useful observation: ``The most frequent packet sizes in the Tor traffic dumps are, in descending order, 1500, −52, −638, 52, 638 and 1150 bytes, accounting for 87.6\% of all Tor packets.''

Hidden Markov Models (HMMs) are first used as a traffic analysis technique in \emph{HMM profiles for network traffic classification} \parencite{Wright:2004p3860}. The primary identification characteristics for use with this algorithm are packet size and inter-packet arrival times. With refinement this algorithm is used with increasing accuracy in \textcite{Wright:2006p322} and \textcite{Dainotti:2008p1435}. HMMs are also used in \textcite{Bernaille:2005p6205} to discover distinguishing characteristics of traffic flows, rather than specifically as a classifier.

@@

@@

\subsection{Comparing Techniques}

It is difficult to say which technique would be most suitable for identifying Tor traffic. In the literature, no particular consensus has been reached and no two papers use the same baseline data set for their technique and so no comparison can be made \parencite{Kim:2007p3867}. Although there have been some comparisons of machine learning algorithms \parencite{Williams:2006p3849,Mohd:2009p6484}, no paper has published their data set due to privacy concerns. It is clear that choosing the right technique will require some experimentation and comparison of the effectiveness of each algorithm.

@@

\begin{figure}[H]
\center \includegraphics[scale=0.7]{network-diagram}
\caption{Network Diagram}
\label{network-diagram}
\end{figure}

\subsection{Methodology}

\subsubsection{Data Collection}

The data collection phase will be conducted by running a series of simulations designed to mimic real world traffic. These simulations will be controlled so that the data captured will only contain Tor traffic in the training and testing data sets, and non Tor encrypted traffic in the data set for determining the false positive rate.

\subsubsection{Algorithm Selection}

Once the data collection phase is complete, a number of different algorithms will be chosen to classify Tor traffic. These will be compared using a number of criteria to rate effectiveness. These will include the accuracy of the matching technique and the number of packets needed to make a confident identification.
